---
title: "miR Enrichment"
author: "Burnet Bioinformatics"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 7
    fig_height: 7
theme: cosmo
---

Source: https://github.com/markziemann/miR-enrichment

## Introduction

The predicted targets of microRNAs are used in functional enrichment analysis to justify the
potential function of microRNAs, but there are some logical problems with this.
In a biological tissue, not all of these targets will be expressed.
Also the enrichment analysis requires a background list which is all the genes that can be
measured.
Not all genes are expressed in a tissue, at least half are silenced, so each enrichment
analysis requires a custom background gene list.
Unfortunately a custom background list is rarely used.
We argue that this causes a dramatic distortion to the results.

## Sample sheet

GSE188236

Control mRNA: SRR16776697, SRR16776698, SRR16776699, SRR16776700, SRR16776701

Case mRNA: SRR16776702, SRR16776703, SRR16776704, SRR16776705, SRR16776706

## Methods

```{r,packages}

suppressPackageStartupMessages({
    library("DESeq2")
    library("gplots")
    library("mitch")
    library("eulerr")
    library("getDEE2")
    library("kableExtra")
})

```

## Import read counts

Importing RNA-seq data

```{r,importdata}

myfiles <-list.files(".",pattern="ke.tsv",recursive=TRUE)

x <- lapply(myfiles,function(x) {
  xx <- read.table(x,header=TRUE,row.names=1)
  xx[,3,drop=FALSE]
})

x <-  do.call(cbind,x)
colnames(x) <- gsub("_est_counts","",colnames(x))

```

Need gene symbols to map to the transcripts.

```{r,geneinfo}

mdat <- getDEE2Metadata("hsapiens")
d <- getDEE2(species="hsapiens",SRRvec="SRR11509477",mdat,outfile="NULL",counts="GeneCounts",legacy=TRUE)
head(d$TxInfo)
txinfo <- d$TxInfo

```

Merge txinfo.

```{r,txinfo}

xm <- merge(x,txinfo,by=0)
xm$GeneID_symbol <- paste(xm$GeneID,xm$GeneSymbol)
xm$Row.names = xm$GeneID = xm$GeneSymbol = xm$TxLength = NULL
xa <- aggregate(. ~ GeneID_symbol,xm,sum)
rownames(xa) <- xa[,1]
xa[,1] = NULL

```

## Differential expression

```{r,de}

xaf <- xa[which(rowMeans(xa)>=10),]
dim(xa) ; dim(xaf)

ss <- data.frame("run"=colnames(xaf),"trt"=c(0,0,0,0,0,1,1,1,1,1))
rownames(ss) <- ss$run

mds <- cmdscale(dist(t(xaf)))
plot(mds,cex=2,col="gray",pch=19)
text(mds, labels=rownames(mds) ,col="black")

colSums(xaf)

dds <- DESeqDataSetFromMatrix(countData = round(xaf) , colData = ss , design = ~ trt )
res <- DESeq(dds)
z <- results(res)
vsd <- vst(dds, blind=FALSE)
zz <- cbind(as.data.frame(z),assay(vsd))
dge <- as.data.frame(zz[order(zz$pvalue),])
head(dge)

ups <- rownames(subset(dge,padj<0.05 & log2FoldChange>0))
dns <- rownames(subset(dge,padj<0.05 & log2FoldChange<0))
lapply(list("UPs"=ups,"DNs"=dns),length)
nrow(dge)

saveRDS(dge,file="GSE188236.Rds")

```

## Session information

For reproducibility.

```{r,session}

sessionInfo()

```
